---
title: "Figure3_Analysis_Sanger_data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library("lmtest")
library("ggplot2")
library("MASS")
library("fitdistrplus")
library("lme4")
library("boot")


setwd("~/MutationFreq_ABL_IC50/Figures/Figure 3")
compicmut=read.csv("Combined_data_frame_IC_Mutprob_Abundance.csv",header=T,stringsAsFactors=F)

```

## Exploration of the predictor variables
First we use a series of plots to explore the data
```{r , echo=TRUE}

x=ggplot(compicmut,aes(Mutation.Probability,IC50))
x+geom_point(aes(size=1))+theme_classic()+theme(axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))

x=ggplot(compicmut,aes(IC50,med_ctr_growth))
x+geom_point(aes(size=1))+theme_classic()+theme(axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))

x=ggplot(compicmut,aes(med_ctr_growth,Mutation.Probability))
x+geom_point(aes(size=1))+theme_classic()+theme(axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))

```

## Abundance_Sanger values are counts so we will use count regression
A poisson distribution fit to the Abundance_Sanger data appears over dispersed, and a negative binomial distribution fits better.

```{r , echo=TRUE}
plot(fitdist(compicmut$Abundance_Sanger,"pois"))#overdispersed
plot(fitdist(compicmut$Abundance_Sanger,"nbinom"))#fit appears better than poisson
```

##Below are summaries for all single variable models, and the best 2 and 3 variable models
```{r echo=TRUE}
# Of IC50 variables, the best single variable model that includes IC50 is below
fit1 <- glm.nb(Abundance_Sanger ~ log10(IC50), data=compicmut)
summary(fit1)
```

```{r echo=TRUE}
#The best single variable model is a model based upon mutation probab
fit2 <- glm.nb(Abundance_Sanger~Mutation.Probability, data=compicmut) 
#Best single variable model
summary(fit2)
```

```{r echo=TRUE}
#Relative growth is not a good predictor of Abundance_Sanger
fit3<-glm.nb(Abundance_Sanger ~ relative_growth, data=compicmut)
summary(fit3)
```

```{r echo=TRUE}
#Median centering does not help growth predict Abundance_Sanger
fit4<-glm.nb(Abundance_Sanger ~ med_ctr_growth, data=compicmut)
summary(fit4)
```

```{r echo=TRUE}
#The best 2 variable model includes a log transformation of growth in the presence of drug(IC50) and mutation probability
fit5<-glm.nb(Abundance_Sanger ~ log10(IC50)+log10(Mutation.Probability), data=compicmut) #Best 2 variable model
summary(fit5)
```

```{r echo=TRUE}
fit7<-glm.nb(Abundance_Sanger ~IC50+log10(Mutation.Probability)+relative_growth, data=compicmut) #Best 3 variable model
summary(fit7)# While AIC decreases here,the decrease is small, the coefficient is negative (which is unlikely), and the pval is not particularly significant. Therefore it seems that this best 3 variable model is likely overfit. I will examine LOOCV, and test set mean squared error for final model selection.
```

```{r}
#Build a data frame for plotting
Data_fit=data.frame(cbind(compicmut,fit5$fitted.values,fit1$fitted.values,fit2$fitted.values))
```

```{r}
#This plots best fit of the IC50 data
x=ggplot(data =Data_fit,aes(fit1.fitted.values,Abundance_Sanger))
x+geom_point(aes(size=1))+theme_classic()+stat_function(fun=function(x)x, geom="line", aes(colour="square",size=1.0),)+theme(axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
```

#Plot the single variable mutation probability model
```{r}
x=ggplot(data =Data_fit,aes(fit2.fitted.values,Abundance_Sanger))
x+geom_point(aes(size=1))+theme_classic()+stat_function(fun=function(x)x, geom="line", aes(colour="square",size=1.0),)+theme(axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
```

#Observed versus predicted for best 2 variable model
```{r}
x=ggplot(data =Data_fit,aes(fit5.fitted.values,Abundance_Sanger))
x+geom_point(aes(size=1))+theme_classic()+stat_function(fun=function(x)x, geom="line", aes(colour="square",size=1.0),)+theme(axis.text.x = element_text(size=15),axis.text.y = element_text(size=15))
```

## Calculate the significance of the model with/without mutation probability
Model is VERY significantly improved by including mutation probability
```{r echo=TRUE}
f5no_mut <- update(fit5, . ~ . -log10(Mutation.Probability) )
anova(fit5, f5no_mut)
```

## Calculate whether a Negative Binomial regression and therefore the second distribution parameter is supported by the data.
```{r echo=TRUE}
fit5pois <- glm(Abundance_Sanger ~ log10(Mutation.Probability)+IC50, family = "poisson", data = compicmut)
pchisq(2 * (logLik(fit5) - logLik(fit5pois)), df = 1, lower.tail = FALSE)
```

##To confirm the choice between the best 2 and 3 variable models, we will use cross validation
```{r echo=TRUE}
record=matrix(nrow=4,ncol=nrow(compicmut))
for (i in 1:nrow(compicmut)){
  train=compicmut[-i,]
  test=compicmut[i,]
  fit1cv=glm.nb(Abundance_Sanger~Mutation.Probability, data=train)#Best 1 variable model
  fit2cv=glm.nb(Abundance_Sanger ~ log10(IC50)+log10(Mutation.Probability), data=train) #Best 2 variable model
  fit3cv=glm.nb(Abundance_Sanger ~ IC50+log10(Mutation.Probability)+relative_growth,data=train) #Best 3 variable model
  record[1,i]=predict(fit1cv, test, type = "response")
  record[2,i]=predict(fit2cv, test, type = "response")
  record[3,i]=predict(fit3cv, test, type = "response")
  pred=predict(fit2cv, test, type = "response", se.fit=T)
  record[4,i]=pred$se.fit
}
MSE=matrix(nrow=3,ncol=1)
#TestMSE
MSE[1,1]=round((sum((record[1,]-compicmut$Abundance_Sanger)^2))/nrow(compicmut),digits=2)
MSE[2,1]=round((sum((record[2,]-compicmut$Abundance_Sanger)^2))/nrow(compicmut),digits = 2)
MSE[3,1]=round((sum((record[3,]-compicmut$Abundance_Sanger)^2))/nrow(compicmut),digits = 2)
MSE=data.frame(cbind(c(1,2,3),as.numeric(as.character(c(MSE[,1]))),c("LOOCV","LOOCV","LOOCV")), stringsAsFactors = F)

colnames(MSE)[1:3]=c("Variable_Number_In_Model","MSE","Type")
MSE=transform(MSE, MSE = as.numeric(MSE))

```

## The LOOCV mean squared error(MSE) of the test observation for the best 1,2,3 variable models
Note that a minimum is seen at 2 variables, and that this is consistent with a model that only includes mutation probability and IC50
```{r}
x=ggplot(data=MSE, aes(x=Variable_Number_In_Model,y=MSE))
x+geom_bar(stat="identity")+theme_classic()
```

```{r}
#Build data frame
binconU=sqrt(sum(Data_fit$Abundance_Sanger)*(Data_fit$Abundance_Sanger/sum(Data_fit$Abundance_Sanger))*((sum(Data_fit$Abundance_Sanger)-Data_fit$Abundance_Sanger)/sum(Data_fit$Abundance_Sanger)))+Data_fit$Abundance_Sanger
binconL=Data_fit$Abundance_Sanger-sqrt(sum(Data_fit$Abundance_Sanger)*(Data_fit$Abundance_Sanger/sum(Data_fit$Abundance_Sanger))*((sum(Data_fit$Abundance_Sanger)-Data_fit$Abundance_Sanger)/sum(Data_fit$Abundance_Sanger)))
Data_fit=data.frame(cbind(compicmut,fit5$fitted.values,fit1$fitted.values,fit2$fitted.values,record[2,],record[4,],binconU,binconL))
colnames(Data_fit)[9:15]=c("Vars_3","Vars_1","Vars_2","CV_2Var_mod","sd","binconL","binconU")
```

#Observed vs Predicted for cross validation
```{r}
plot(Data_fit$Abundance_Sanger,Data_fit$CV_2Var_mod,xlim=c(1,250),ylim=c(1,250),
    pch=19,cex=2.5,cex.axis=2,cex.lab=1.5, xlab="Measured Counts", ylab="Predicted Counts")
# hack: we draw arrows but with very special "arrowheads"
arrows(Data_fit$Abundance_Sanger,Data_fit$CV_2Var_mod-Data_fit$sd,Data_fit$Abundance_Sanger,Data_fit$CV_2Var_mod+Data_fit$sd, length=0.09, angle=90,lwd=3, code=3)
arrows(Data_fit$binconL,Data_fit$CV_2Var_mod,Data_fit$binconU,Data_fit$CV_2Var_mod, length=0.09, angle=90,lwd=3, code=3)
abline(a=0,b=1,col="red",lwd=3)
m=lm(Vars_2~Abundance_Sanger+0,data=Data_fit)
summary(m)
```


